---
title: "Election of 2014 in Värmlands län"
author: "Martin"
date: '2018-08-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(sf)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
#load_data, results from the election:

valres_2014 <- read_csv2("https://data.val.se/val/val2014/statistik/2014_riksdagsval_per_valdistrikt.skv",
                         locale=locale("sv",encoding="ISO-8859-1"), quote ="")

#map as shapefile and filter it down to my region (region code starts with 17):
shp_varmland <- read_sf("data/valgeografi_valdistrikt_2014/valgeografi_valdistrikt.shp") %>% 
  filter(str_detect(VD, "^17"))

#filter the results_file to fit my region and select the interesting colums
valres_varmland_2014 <- valres_2014 %>% 
  select(LAN,
         KOM,
         VALDIST,
         RIKSDAGSVALKRETS,
         KOMMUN,
         Kommunvalkrets,
         Valdistrikt,
         `M proc`,
         `C proc`,
         `FP proc`,
         `KD proc`,
         `S proc`,
         `V proc`,
         `MP proc`,
         `SD proc`,
         VDT) %>% 
  filter(RIKSDAGSVALKRETS=="Värmlands län",
         VALDIST!="00" & VALDIST!="01" & VALDIST!="02") %>% 
  rename(Moderaterna=`M proc`,
         Centerpartiet=`C proc`,
         Liberalerna=`FP proc`,
         Kristdemokraterna=`KD proc`,
         Socialdemokraterna=`S proc`,
         Vänsterpartiet=`V proc`,
         Miljöpartiet=`MP proc`,
         Sverigedemokraterna=`SD proc`)
```

## The 2014 Riksdag election

This is a presentation of the 2014 election to the national parliament, Riksdagen, in Sweden and how the people of the county of Värmland voted. Värmland is divided in 176 election districts with approximatly 1.000 voters in each. This is the voter's turnout for every district:

```{r valres_varmland_2014, echo=F}

library(ggplot2)
library(gghighlight)

ggplot(valres_varmland_2014, aes(x=VDT, y=KOMMUN), size=2) +
  geom_point()+
  theme_minimal()+
  gghighlight(VDT==max(VDT), label_key=Valdistrikt, use_group_by=F)+
  theme(axis.line.y=element_blank(),
        panel.grid.major.y = element_blank())+
  labs(title="Voter turnout in voting districts",
       x="Voter turnout in percent",
       y="",
       caption="Source: Val.se")
  
```

## Party support in Värmland

There are eight parties in the Swedish parliament. These maps show the support for each of them in the last election.

```{r map, echo=FALSE, warning=FALSE}
#load libraries and data

library(ggplot2)
library(sf)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(DT)
library(gghighlight)
library(forcats)

valres_2014 <- read_csv2("https://data.val.se/val/val2014/statistik/2014_riksdagsval_per_valdistrikt.skv",
                         locale=locale("sv",encoding="ISO-8859-1"), quote ="")

#map as shapefile and filter it down to my region (region code starts with 17):
shp_varmland <- read_sf("data/valgeografi_valdistrikt_2014/valgeografi_valdistrikt.shp") %>% 
  filter(str_detect(VD, "^17"))

#filter the results_file to fit my region and select the interesting colums
valres_varmland_2014 <- valres_2014 %>% 
  select(LAN,
         KOM,
         VALDIST,
         RIKSDAGSVALKRETS,
         KOMMUN,
         Kommunvalkrets,
         Valdistrikt,
         `M proc`,
         `C proc`,
         `FP proc`,
         `KD proc`,
         `S proc`,
         `V proc`,
         `MP proc`,
         `SD proc`,
         VDT) %>% 
  filter(RIKSDAGSVALKRETS=="Värmlands län",
         VALDIST!="00" & VALDIST!="01" & VALDIST!="02") %>% 
  rename(Moderaterna=`M proc`,
         Centerpartiet=`C proc`,
         Liberalerna=`FP proc`,
         Kristdemokraterna=`KD proc`,
         Socialdemokraterna=`S proc`,
         Vänsterpartiet=`V proc`,
         Miljöpartiet=`MP proc`,
         Sverigedemokraterna=`SD proc`)

# join the two datasets
map_varmland_2014 <- left_join(shp_varmland, valres_varmland_2014, by=c("VD_NAMN"="Valdistrikt"))

# make wide dataset long, for visualization purposes

map_varmland_2014_long <- gather(map_varmland_2014, party, proc_of_votes, Moderaterna:Sverigedemokraterna, factor_key = TRUE)

#start visualising

ggplot(map_varmland_2014_long) +
  geom_sf(aes(fill=proc_of_votes), color="white", size=.1) +
  theme_void()+
  theme(panel.grid.major = element_line(colour="transparent"),
        legend.position = "bottom") +
  facet_wrap(~party, ncol=4) +
  scale_fill_distiller(palette="Oranges", direction=1, name="Percent of votes") +
  labs(title="How Värmland Votes: Party Support in the 2014 Riksdag election",
       caption="Source: Val.se")
```


